The Driving Forces Behind API Evolution

### 2015-05-20 thesis/paper.tex
Changed heading names
Fixed typo
Typo in RQs
Updates to comments in discussion

\subsection{Research Questions} \label{rqs}
\begin{description}
\item[\textbf{RQ1}] What are the driving forces to API evoltion?
\item[\textbf{RQ2}] To what extent can we reverse engineer API change decisions?

@@

Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman90sview}, known as Lehman's laws. They are presented in table \ref{table:expl_laws}. The laws are said to only apply to E-type systems, which are programs that solve problems or address applications in the real world.

@@

\subsection{RQ1 Results} \label{results_rq1}

@@

\subsection{RQ2 Results}
\noindent
RQ2: \textbf{To what extent can we reverse engineer API change decisions?}
\smallskip

@@

\subsection{RQ1 Answer}  
Our results show that the major driving force for change in APIs is the push (!!!reword) for functionality. The classification of the types of software evolution and software maintenance showed that enhancive changes which either aimed to add brand-new (!!!reword) modules or adding functionality to existing modules were the most common. We also think that it is interesting to note that a majority of the changes which did not affect functionality, such as the updative changes, were spurred because of previous changes which were related to functionality. Following this line of thought, the hunger (!!!reword?) for functionality can be seen as a catalyst for other driving forces to occur (!!!reword). The view of considering functionality to be a major driver for change is supported by the analysis with Lehman's laws. In the law of Continuing Change, Lehman concludes that ``an E-type system must be continually adapted or it becomes progressively less satisfactory". The results of our work supports this view, as the changes did not only add functionality, but it was also found that the application developers most major concern were updates to functionality.

In relation to this, previous work has identified that refactorings make up a substantial part of the changes APIs go through (!!!FIND PAPERS TO CITE HERE). These results are not in line with ours, since our results show that groomative changes are uncommon and that changes related to quality are not primary drivers of change and that they rather are ripple-effects of changes to functionality.

Our results found that seven out of eight of Lehman's laws do apply to API evolution. The law of Conservation of Organisational Stability was found not to apply, since changes to APIs are often discouraged (!!!reword) by the fear that it will disrupt the applications using the API. This inertia to change in API evolution is supported by other research, which also conclude that the negative effect that changes to APIs may have is a major factor to be considered \cite{google_talk, henning2007api, mcdonnell2013empirical, robbes2012developers}.

Based on our results, we form the conclusion that the hunger (!!!reword) for functionality and the fear (!!!reword) of changing are two opposing forces that cause updates to APIs to be infrequent. In the evolution of APIs, the negative effects of changes are greater than in other types of systems, which causes the the law of Conservation of Organisational Stability to not apply (!!!reword). This occurring has been considered by Lehman, as he noted that ``the laws are not immutable, since they arise from the habits and practices of humans" \cite{lehman1980programs}. The underlying reason for this is that the laws are influenced by conscious thought processes, based on human understanding \cite{lehman1980understanding}, which in the case of APIs is the conscious decision to limit updates as much as possible.

\subsection{RQ2 Answer}
We conclude that we were able to reverse engineer the change decisions to a large extent. This is clear from the results, which showed that 85\% of the changes were reverse engineered correctly. The kappa value of 0.812 further supports this. We interpret that the significance of this result suggests that reverse engineering of change decisions can be done in situations where a perfect result is not required. In cases where a near-perfect result is enough, we do, however, believe that reverse engineering change decisions is worthwhile (!!!this sentence weird). This line of thought is supported by our results, which, in summary, show that we are able to reverse engineer the most major trends with very high accuracy, but that certain nuances of the changes can be misinterpreted.

@@

By analysing the API based on Lehman's laws, we were able to find that the law of Conservation of Organisational Stability did not apply to the API examined in this study. Our results suggest that the opposite of what the law says in its current form is closer to the truth (!!!reword) for APIs. Since we only studied one API, we cannot safely refute the law or propose a new formulation of it. We do, however, believe that academia should be wary of assuming (!!!reword) that this law is true for APIs.

\subsection{Implications for Practitioners}
The changes introduced to the API are mainly based on the expertise and experiences of the API architects at the case company. There has not been a scientific approach to maintaining the system and the knowledge from academia has not directly affected the way that the API's evolution has been managed. With our work, we hope to provide a new type of insight (!!!reword) into how the API is maintained. Our analysis of the API based on Lehman's laws should further assure that the case company is moving the evolution of the API in the right direction. With our research, we leave a challenge (!!!reword) for the case company in how to tackle challenges related to the law of Conservation of Organisational Stability. Determining if the benefits of continuously and frequently updating the API outweigh the negative effects that these updates could have on the application developers, we leave up to the company (!!!reword so that it doesnt sound negative)

One way of finding the solution to this problem would be by including the application developers in the feedback loop. Since the application developers are on the receiving end of the changes, we believe that their needs have to be assessed before the problem can be solved. With our study we identified both that the law of Feedback System is of great importance to the case company and that it is one of the main aspects of the development of their API which they want to improve (!!!reword). One challenge of software evolution is that the activities carried out are often intangible and non-objective \cite{chapin2001types, lientz1980software}. Based on this, we propose that our study which has granulated the activites is used as a basis for communicating the current stage of the evolution of their API. We hope that this classification can be used in its current form, to help quantify the activities and to provide an overview of them. We also hope that the approach we used can be used to predict which types of activites should be emphasised in the future evolution of the API. With our study we have also demonstrated that the classification of changes is achievable, which, in turn, should encourage the case company to follow suit (!!!casual).

>>>>

\subsection{Research Questions} \label{rqs}
\begin{description}
\item[\textbf{RQ1}] What are the driving forces of API evolution?
\item[\textbf{RQ2}] To what extent can we reverse engineer API change decisions?

@@

Software evolution theory suggests that changes to software must obey one of eight laws \cite{lehman90sview}, known as Lehman's laws. They are presented in Table \ref{table:expl_laws}. The laws are said to only apply to E-type systems, which are programs that solve problems or address applications in the real world.

@@

\subsection{The Driving Forces of API Evolution} \label{results_rq1}

@@

\subsection{Reverse Engineering API Change Decisions}
\noindent
RQ2: \textbf{To what extent can we reverse engineer API change decisions?}
\smallskip

@@

\subsection{RQ1 Answer}  
Our results show that the major driving force for change in APIs is the call for increased functionality. The classification of the types of software evolution and software maintenance showed that enhancive changes which either aimed to add modules or new functionality to existing modules were the most common. We also think that it is interesting to note that a majority of the changes which did not affect functionality, such as the updative changes, were spurred because of previous changes that were related to functionality. Following this line of thought, the desire for new functionality can be seen as a catalyst that allows further driving forces to occur (!!!better?). The view of considering functionality to be a major driver for change is supported by the analysis with Lehman's laws. In the law of Continuing Change, Lehman concludes that ``an E-type system must be continually adapted or it becomes progressively less satisfactory". The results of our work supports this view, as the changes did not only add functionality, but it was also found that the application developers most major concern were updates to functionality.

In relation to this, previous work has identified that refactorings make up a substantial part of the changes APIs undergo (!!!FIND PAPERS TO CITE HERE). These results are not in line with ours, since our results show that groomative changes are uncommon and that changes related to quality are not primary drivers of change, rather, they are ripple-effects of changes to functionality \cite{robbes2012developers}.

Our results found that seven out of eight of Lehman's laws do apply to API evolution. The law of Conservation of Organisational Stability was found not to apply, since changes to APIs are unfavourable due to the potential disruption the applications using the API could face. This inertia to change in API evolution is supported by other research, which also conclude that the negative effect that changes to APIs may have is a major factor to be considered \cite{google_talk, henning2007api, mcdonnell2013empirical, robbes2012developers}.

Based on our results, we form the conclusion that the desire for functionality and the reluctance of changing are two opposing forces that cause updates to APIs to be infrequent. In the evolution of APIs, the negative effects of changes are greater than in other types of systems, which causes the the law of Conservation of Organisational Stability to not apply (!!! which means the law of Conservation of Organisational Stability is inappropriate for APIs). This occurring has been considered by Lehman, as he noted that ``the laws are not immutable, since they arise from the habits and practices of humans" \cite{lehman1980programs}. The underlying reason for this is that the laws are influenced by conscious thought processes, based on human understanding \cite{lehman1980understanding}, which in the case of APIs is the conscious decision to limit updates as much as possible.

\subsection{RQ2 Answer}
We conclude that we were able to reverse engineer the change decisions to a large extent. This is clear from the results, which showed that 85\% of the changes were reverse engineered correctly. The kappa value of 0.812 further supports this. We interpret that the significance of this result suggests that reverse engineering of change decisions can be done in situations where a perfect result is not required. In cases where a near-perfect result is enough, we do, however, believe that reverse engineering change decisions is worthwhile (!!!this sentence weird - yeah). This line of thought is supported by our results, which, in summary, show that we are able to reverse engineer the most major trends with very high accuracy, but that certain nuances of the changes can be misinterpreted.

@@

By analysing the API based on Lehman's laws, we were able to find that the law of Conservation of Organisational Stability did not apply to the API examined in this study. Our results suggest that the opposite of what the law suggests, in its current form, is more applicable for APIs. Since only a single API was studied, we cannot safely refute the law or propose a new formulation of it. Academia should therefore be wary of assuming that this law is true for APIs.

\subsection{Implications for Practitioners}
The changes introduced to the API are mainly based on the expertise and experiences of the API architects at the case company. There has not been a scientific approach to maintaining the system and the knowledge from academia has not directly affected the way that the API's evolution has been managed. With our work, we hope to provide further insight into how the API is maintained. Our analysis of the API based on Lehman's laws should further assure that the case company is moving the evolution of the API in the right direction. Our research proposes a challenge for the case company in how to tackle challenges related to the law of Conservation of Organisational Stability. Determining if the benefits of continuously and frequently updating the API outweigh the side-effects that these updates could have on the application developers, is something that the case company will need to consider.

One way of finding the solution to this problem would be by including the application developers in the feedback loop. Since the application developers are on the receiving end of the changes, we believe that their needs have to be assessed before the problem can be solved. With our study we identified both that the law of Feedback System is of great importance to the case company and that it is one of the key areas of API development that the case company wishes to improve. One challenge of the challenges posed by software evolution is that the activities carried out are often intangible and non-objective \cite{chapin2001types, lientz1980software}. Based on this, we propose that our study which has granulated the activities is used as a basis for communicating the current stage of the evolution of their API. We hope that this classification can be used in its current form, to help quantify the activities and to provide an overview of them. We also hope that the approach we used can be used to predict which types of activities should be emphasised in the future evolution of the API. With our study we have also demonstrated that the classification of changes is achievable, which, in turn, should encourage the case company adopt this method of classifying change.
